---
title: "Prediction_for_inhomogenous_Poisson_process"
author: "William Denault"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to Poisson-fSuSiE}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r , include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 5,
  fig.height = 3,
  fig.align = "center",
  fig.cap = "&nbsp;",
  dpi = 175
  )
```


```{r setup, include=FALSE}
rm(list = ls())
library(mvPoisVA)
library(fsusieR)
library(susieR)
library(ebnm)
set.seed(1)
'%!in%' <- function(x,y)!('%in%'(x,y))
N =700
data(N3finemapping)
Z <- matrix (runif(N*200, min=0,max=1.5), nrow =N, ncol = 200)
mysd=0.2

 
```






Here we assume that we observe some covariates that affect the rate of the Poisson process.  We want to build a prediction model that 

```{r}

 
lev_res =7
count.data  <- list()
N_effect=30
lf <-  list()
for ( j in 1:N_effect){
  lf[[j]]<- rep(0.001, 2^lev_res )
  start= sample(1:(2^lev_res-11), size=1)
lf[[j]][start:(start+10)] <-1
 lf[[j]]=  (lf[[j]])
}




true_pos <- sample(1:ncol(genotype), replace = FALSE,size=2)

plot (lf[[1]], type="l", main=paste ( "effect of SNP",true_pos[1]))

plot (lf[[2]], type="l", main=paste ( "effect of SNP",true_pos[2]))
```



Simulating the data

```{r}
 
predictor <-rep (0, length(lf[[1]] ))
count.data  <- list()
intensity= list()
for ( i in 1:N)
{

  predictor <-rep (0, length(lf[[1]] ))

  for ( j in 1:N_effect){
    predictor <-predictor + Z[i, j]*lf[[j]] 
  }
  intensity[[i]]=predictor 
  predictor <- exp( predictor+ rnorm(  length(lf[[1]]), sd=mysd))

  
  count.data [[i]] <-   rpois(n= length(lf[[1]]) ,
                              lambda =predictor  )

}
count.data <- do.call(rbind, count.data)


Y <- count.data

int= do.call(rbind, intensity)
image(log(Y))
hist(Y, nclass=100, main="Histogram of the count data")
```
 
```{r}

res= acc_Pois_fSuSiE2(Y=Y ,Z=Z)
```
We can now check the performance of the model by comparing the estimated underlying intensity
vs the true intensity. 

```{r}

plot(Z%*%res$EBmvFR.obj$fitted_func, int)
lm_fit = lm(c(int)~  c( Z%*%res$EBmvFR.obj$fitted_func))

summary(lm_fit)

```
We can also compare the predicted count data vs the true count data. 
In green the exp of the predicted intensity vs the true count data.
in black the exp of the true intensity vs the true count data.
```{r}


 
plot(exp(Z%*%res$EBmvFR.obj$fitted_func), Y, col="green")
points( exp(int) , (Y))

```

Out of sample prediction:

generate the data

```{r}
Z2 <- matrix (runif(N*200, min=0,max=1.5), nrow =N, ncol = 200)




predictor <-rep (0, length(lf[[1]] ))
count.data2  <- list()
intensity2= list()
for ( i in 1:N)
{

  predictor <-rep (0, length(lf[[1]] ))

  for ( j in 1:N_effect){
    predictor <-predictor + Z2[i, j]*lf[[j]] 
  }
  intensity2[[i]]=predictor 
  predictor <- exp( predictor+ rnorm(  length(lf[[1]]), sd=mysd))

  
  count.data2 [[i]] <-   rpois(n= length(lf[[1]]) ,
                              lambda =predictor  )

}
count.data2 <- do.call(rbind, count.data2)


Y2 <- count.data2
int2= do.call(rbind, intensity2)
```


```{r}

plot(Z2%*%res$EBmvFR.obj$fitted_func, int2)
lm_fit = lm(c(int2)~  c( Z2%*%res$EBmvFR.obj$fitted_func))

summary(lm_fit)

```


```{r}
 
plot(exp(Z2%*%res$EBmvFR.obj$fitted_func), Y2, col="green")
points( exp(int2) , (Y2))
```


